DeepSphere: Efficient spherical Convolutional Neural Network with HEALPix sampling for cosmological applications

### 2019-03-04 rebuttal.tex

\begin{mdframed}[style=comment]
Similarly as (2), the convolution of the signal $f$ by a kernel $g$ is the scalar product of $f$ with translated versions $T_ig$ of the kernel $g$.

- I think this should be "As with equation (2)" or "Similarly, as in equation (2)" but I can't tell which. Also, it would be really useful if you referred to equations by "equation (x)" rather than just "(x)" throughout - much easier to read that way.
\end{mdframed}
% \nati{I do not know what we should do.}
Ideally we would replace equation references such as "(1)" with "Equation (1)", but given the large number of equation references, we would prefer to stick with a short-form to avoid lengthening the paper further.
% \mdeff{Agree to not introduce more words.}

\begin{mdframed}[style=comment]
The 1-neighborhood of a node is the set of nodes that are directly connected to it. The k-neighborhood of a node is the set of nodes that are connected to it through paths of length k.

- I'm a little confused here, I think the neighbourhood is every pixel which is closer than the distance of some pixel. The value of k would then be the length of the set of ordered distances for all pixels up to this pixel. Is that correct? For example, take a 3x3 square (on Euclidean space for simplicity), if the central pixel has coordinates (0, 0) then the 1-neighbourhood would be all pixels at a length of 1 pixel or less, i.e. (-1, 0), (+1, 0), (0, -1) and (0, +1) who are all at the same distance so the set of distances is {1} pixel and the length of the set is 1. The 2-neighbourhood would then be all pixels which were a distance $\sqrt{2}$ pixels or less away, i.e. (-1, 0), (+1, 0), (0, -1), (0, +1), (-1, -1), (-1, +1), (+1, -1), (+1, +1) so the set of distances would be {1, $\sqrt{2}$} pixels and the length of the set is 2. Is this correct? If so k is not a length. If this is not correct, I think I misunderstand something and so this should be updated in the text.
\end{mdframed}

The neighborhood is defined in the general context of graphs, where vertices don't necessarily have coordinates.
One way to define a distance between two vertices is to count the number of edges in the shortest path between them.
We clarified this point.

@@

\begin{mdframed}[style=comment]
Similarly, each line of the matrix $\sum_{k=0}^K\theta_k\bf{L}^k$ defines an irregular patch of radius $K$.

 - What do you mean by "line"? Is it a row or a column or something else which I've misunderstood?
\end{mdframed}
We mean ``column'' and have updated the manuscript.

@@

\begin{mdframed}[style=comment]
Given a map $\bf{x}\in\mathbb{R}^{N_{pix}}$, pooling defines $\bf{y}\in\mathbb{R}^{N'_{pix}}$ such as

- It should be "such that" not "such as". "Given a map $\bf{x}\in\mathbb{R}^{N_{pix}}$, pooling defines $\bf{y}\in\mathbb{R}^{N'_{pix}}$ such that"
\end{mdframed}
Thank you for spotting this mistake. We have corrected the manuscript.

\begin{mdframed}[style=comment]
... where $f$ is a function which operates on sets (possibly of varying sizes) and $N_{pix}/N'_{pix}$ is the down-sampling factor, which for HEALPix is $|\mathcal{C}(i)| = N_{pix}/N'_{pix} = (N_{side} / N'_{side})^2 = 4^p$.

- Can you explain what $p$ is and it would probably be worth splitting this equation into a multi-line align environment for clarity.
\end{mdframed}
$p$ was introduced to highlight the fact that the coarsening factor is a power of 4 by construction of HEALPix.
It is defined as $p=\log_2(N_{side} / N'_{side})$.
We have updated the manuscript with this clarification and with an equation environment.

### 2019-03-04 scnn.tex
comment 2.14 (details) part 2

Finally, the graph convolution can be interpreted in the vertex domain as a scalar product with localizations $\T_i h$ of the kernel $h$. Indeed, the result of the convolution of the signal $\f$ with the kernel $h$ is
\begin{equation} \label{eqn:graph_convolution_spatial}
	(h(\L) \f)_i = \langle \T_i h(\L), \f \rangle = \langle h(\L) \b \delta^i, \f \rangle.
\end{equation}
To make the parallel with the classical 1D convolution, let $f, g\in \mathbb{Z} \rightarrow \mathbb{R}$ be two 1D discrete signals. Their convolution can be written in the same form as \eqnref{graph_convolution_spatial}:
\begin{equation*}
	(f \ast g) [i] = \sum_{j=-\infty}^\infty f[j] g[i-j] = \langle T_i g,  f \rangle,
\end{equation*}
where $T_i g[j] = g[i-j]$ is, up to a flip ($g[i-j]$ instead of $g[i+j]$), a translation operator.
Similarly as \eqnref{graph_convolution_spatial}, the convolution of the signal $f$ by a kernel $g$ is the scalar product of $f$ with translated versions $T_i g$ of the kernel $g$.
Additionally, it turns out that the localization operator $\T_i h$ is a generalization of the translation operator on graphs. In the particular case where the Laplacian matrix $\L$ is circulant, $\T_i h$ is a translated version of $\T_j h$ for all $i, j$ and both convolutions are equivalent. We refer the reader to \citep[Sec 2.2]{perraudin2017stationary} for a detailed discussion of the connection between translation $T_i$ and localization $\T_i$.

@@

Fortunately, both of these computational issues are overcome by defining the convolution kernel $h$ as a polynomial $h_\theta(\lambda) = \sum_{k=0}^K \theta_k \lambda^k$ of degree $K$ parametrised by $K+1$ coefficients $\theta$.
The filtering operation \eqnref{graph_convolution_fourier} becomes
\begin{equation} \label{eqn:graph_convolution_monomial}
	h_\theta(\L) \f =  \U \left(\sum_{k=0}^K \theta_k \bLambda^k \right) \U\trans \f = \sum_{k=0}^K \theta_k \L^k \f,
\end{equation}
where $\L^k$ captures $k$-neighborhoods.\footnote{In the context of graph, one way to define the distance between two vertices is to count the number of edges in the shortest path between these two vertices. Hence, we define the 1-neighborhood of a vertex as the set of vertices that are directly connected to it. The $k$-neighborhood of a vertex is the set of vertices that are connected to it through paths of length $k$.}
The entry $(\L^k)_{ij}$ is the sum of all weighted paths of length $k$ between vertices $v_i$ and $v_j$, where the weight of a path is the multiplication of all the edge weights on the path.
Hence, it is non-zero if and only if vertices $v_i$ and $v_j$ are connected by at least one path of length $k$.
Filtering with a polynomial convolution kernel can thus be interpreted in the pixel (vertex) domain as a weighted linear combination of neighboring pixel values.
In the classical setting, convolutions are also weighted local sums.
The weights are however given by the filter coefficients only, and there is one coefficient per pixel in the patch.
In the graph setting, the weights are determined by the filter coefficients $\theta$ and the Laplacian $\L$, and there is one coefficient per neighborhood, not vertex.
That defines radial filters, values of which depend only on the distance to the center, and not on the direction.
% It is required as there is no way to order and match vertices from the same neighborhood.
% The number of vertices in a neighborhood most often even vary from vertex to vertex.
While it may seem odd to restrict filters to be 1D while the sphere is 2D, radial filters are direction-less and result in rotation equivariant convolutions.
We note that restricting the graph convolutional kernel to a polynomial is similar to restricting the classical Euclidean convolution to a fixed-size patch.
Similarly, each column of the matrix $\sum_{k=0}^K \theta_k \L^k$ defines an irregular patch of radius $K$.
Hence, filters designed as polynomials of the Laplacian have local support in the vertex domain.

@@

Neural networks are constructed as stacks of layers which sequentially transform the data from its raw representation to some predictions.
The general DeepSphere architecture, pictured in \figref{architecture}, is composed of many layers. The convolutional part, the head of the neural network, is composed of graph convolutional layers ($GC$), pooling layers ($P$), and batch normalization layers ($BN$). The tail is composed of multiple fully connected layers ($FC$) followed by an optional softmax layer ($SM$) if the network is used for discrete classification.
A non-linear function $\sigma(\cdot)$ is applied after every linear $GC$ and $FC$ layer, except for the last $FC$ layer. That operation is point-wise, i.e., $y_{ij} = \sigma(x_{ij})$ and $y_i = \sigma(x_i)$ for matrices $\X, \Y$ and vectors $\x, \y$. The rectified linear unit (ReLU) $\sigma(\cdot) = \max(\cdot, 0)$ is a common choice and is the one we adopted for this contribution.

@@

The layer $FC: \R^{F_{in}} \rightarrow \R^{F_{out}}$ is defined as
\begin{equation} \label{eqn:fc_layer}
	\y = FC(\x) = \b W \x + \b b ,
\end{equation}
where $\W \in \R^{F_{out} \times F_{in}}$ and $\b b \in \R^{F_{out}}$ are the parameters to be learned.
Note that the output $\Y \in \R^{N_{pix} \times F_{out}}$ of the last $GC$ (or the output $\Y \in \R^{N_{stat} \times F_{out}}$ of the statistical layer), is vectorized as $\x = \vect(\X) \in \R^{F_{in}}$ before being fed to the first $FC$, where $F_{in} = N_{pix} \times F_{out}$ (or $N_{stat} \times F_{out}$).

The softmax layer is the last layer in a neural network engineered for classification. Given the output $\x \in \R^{N_{classes}}$ of the last $FC$ (or statistical layer), called the logits in the deep learning literature, the softmax layer outputs $\y = SM(\x)$ such that
\begin{equation*}
	y_i = \frac{\exp(x_i)}{\sum_{j=1}^{N_{classes}} \exp(x_j)} \in [0, 1], \ \forall i \in [N_{classes}],
\end{equation*}
where $N_{classes}$ is the number of classes to discriminate.
Thanks to the softmax, the output $\y \in \R^{N_{classes}}$ of a neural network is a discretised conditional distribution for the class given the data and trained parameters.
That is, $y_i$ is the confidence of the network that the input sample belongs to class $i$.
% Note that in general, this probability distribution does not correspond to the true posterior distribution $p(y|x)$.
This last layer is actually normalizing $\x$ into $\y$ such that $\| \y \|_1 = \sum_i y_i = 1$.

@@

The set of input pixels which influence the value of an output pixel forms a \textit{receptive field}.
That field is isotropic and local, i.e., it forms a disk centered around the output pixel.
The radius of this filter is influenced by the polynomial order $K$ of $GC$ layers (setting how far the convolution operation looks around), and the down-sampling factor of $P$ layers, if any.
This radius should be large enough to capture statistics of interest.
For example, a partial sky observation can provide only limited information of cosmological relevance.
On the other hand, looking at the whole sky is often superfluous and waste computations, as most interactions are local.
A data and task dependent trade-off is to be found.
Note here that while global predictions use, by definition, the whole sky, dense predictions are not necessarily local.

>>>>

Finally, the graph convolution can be interpreted in the vertex domain as a scalar product with localizations $\T_i h$ of the kernel $h$. Indeed, the result of the convolution of the signal $\f$ with the kernel $h$ is
\begin{equation} \label{eqn:graph_convolution_spatial}
	(h(\L) \f)_i = \langle \T_i h(\L), \f \rangle = \langle h(\L) \b \delta^i, \f \rangle.
\end{equation}
To make the parallel with the classical 1D convolution, let $f, g\in \mathbb{Z} \rightarrow \mathbb{R}$ be two 1D discrete signals. Their convolution can be written in the same form as \eqnref{graph_convolution_spatial}:
\begin{equation*}
	(f \ast g) [i] = \sum_{j=-\infty}^\infty f[j] g[i-j] = \langle T_i g,  f \rangle,
\end{equation*}
where $T_i g[j] = g[i-j]$ is, up to a flip (i.e., $g[i-j]$ instead of $g[i+j]$), a translation operator.
Similarly as \eqnref{graph_convolution_spatial}, the convolution of the signal $f$ by a kernel $g$ is the scalar product of $f$ with translated versions $T_i g$ of the kernel $g$.
Additionally, it turns out that the localization operator $\T_i h$ is a generalization of the translation operator on graphs. In the particular case where the Laplacian matrix $\L$ is circulant, $\T_i h$ is a translated version of $\T_j h$ for all $i, j$ and both convolutions are equivalent. We refer the reader to \citep[Sec 2.2]{perraudin2017stationary} for a detailed discussion of the connection between translation $T_i$ and localization $\T_i$.

@@

Fortunately, both of these computational issues are overcome by defining the convolution kernel $h$ as a polynomial $h_\theta(\lambda) = \sum_{k=0}^K \theta_k \lambda^k$ of degree $K$ parametrised by $K+1$ coefficients $\theta$.
The filtering operation \eqnref{graph_convolution_fourier} becomes
\begin{equation} \label{eqn:graph_convolution_monomial}
	h_\theta(\L) \f =  \U \left(\sum_{k=0}^K \theta_k \bLambda^k \right) \U\trans \f = \sum_{k=0}^K \theta_k \L^k \f,
\end{equation}
where $\L^k$ captures $k$-neighborhoods.
The entry $(\L^k)_{ij}$ is the sum of all weighted paths of length $k$ between vertices $v_i$ and $v_j$, where the weight of a path is the multiplication of all the edge weights on the path.
Hence, it is non-zero if and only if vertices $v_i$ and $v_j$ are connected by at least one path of length $k$.\footnote{The length of a path between two vertices defines a distance on the graph.}
Filtering with a polynomial convolution kernel can thus be interpreted in the pixel (vertex) domain as a weighted linear combination of neighboring pixel values.
In the classical setting, convolutions are also weighted local sums.
The weights are however given by the filter coefficients only, and there is one coefficient per pixel in the patch.
In the graph setting, the weights are determined by the filter coefficients $\theta$ and the Laplacian $\L$, and there is one coefficient per neighborhood, not vertex.
That defines radial filters, values of which depend only on the distance to the center, and not on the direction.
% It is required as there is no way to order and match vertices from the same neighborhood.
% The number of vertices in a neighborhood most often even vary from vertex to vertex.
While it may seem odd to restrict filters to be 1D while the sphere is 2D, radial filters are direction-less and result in rotation equivariant convolutions.
We note that restricting the graph convolutional kernel to a polynomial is similar to restricting the classical Euclidean convolution to a fixed-size patch.
Similarly, each column of the matrix $\sum_{k=0}^K \theta_k \L^k$ defines an irregular patch of radius $K$.
Hence, filters designed as polynomials of the Laplacian have local support in the vertex domain.

@@

Neural networks are constructed as stacks of layers which sequentially transform the data from its raw representation to some predictions.
The general DeepSphere architecture, pictured in \figref{architecture}, is composed of many layers. The convolutional part, the head of the neural network, is composed of graph convolutional layers ($GC$), pooling layers ($P$), and batch normalization layers ($BN$). The tail is composed of multiple fully connected layers ($FC$) followed by an optional softmax layer ($SM$) if the network is used for discrete classification.
A non-linear function $\sigma(\cdot)$ is applied after every linear $GC$ and $FC$ layer, except for the last $FC$ layer.
That operation is point-wise, i.e., $y_{ij} = \sigma(x_{ij})$ and $y_i = \sigma(x_i)$ for matrices $\X, \Y$ and vectors $\x, \y$.
The rectified linear unit (ReLU) $\sigma(\cdot) = \max(\cdot, 0)$ is a common choice, and is the one we adopted in this contribution.

@@

The layer $FC: \R^{F_{in}} \rightarrow \R^{F_{out}}$ is defined as
\begin{equation} \label{eqn:fc_layer}
	\y = FC(\x) = \b W \x + \b b ,
\end{equation}
where $\W \in \R^{F_{out} \times F_{in}}$ and $\b b \in \R^{F_{out}}$ are the parameters to be learned.
Note that the output $\Y \in \R^{N_{pix} \times F_{out}}$ of the last $GC$ is vectorized as $\x = \vect(\X) \in \R^{F_{in}}$ before being fed to the first $FC$, where $F_{in} = N_{pix} \times F_{out}$.

The softmax layer is the last layer in a neural network engineered for classification.
Given the output $\x \in \R^{N_{classes}}$ of the last $FC$, called the logits in the deep learning literature, the softmax layer outputs $\y = SM(\x)$ such that
\begin{equation*}
	y_i = \frac{\exp(x_i)}{\sum_{j=1}^{N_{classes}} \exp(x_j)} \in [0, 1], \ \forall i \in [N_{classes}],
\end{equation*}
where $N_{classes}$ is the number of classes to discriminate.
Thanks to the softmax, the output $\y \in \R^{N_{classes}}$ of a neural network is a discretised conditional distribution for the class given the data and trained parameters.
That is, $y_i$ is the confidence of the network that the input sample belongs to class $i$.
% Note that in general, this probability distribution does not correspond to the true posterior distribution $p(y|x)$.
This last layer is actually normalizing $\x$ into $\y$ such that $\| \y \|_1 = \sum_i y_i = 1$.

@@

The set of input pixels which influence the value of an output pixel forms a \textit{receptive field}.
That field is isotropic and local, i.e., it forms a disk centered around the output pixel.
The field's radius is influenced by the polynomial order $K$ of $GC$ layers (setting how far the convolution operation looks around), and the down-sampling factor of $P$ layers, if any.
This radius should be large enough to capture statistics of interest.
For example, a partial sky observation can provide only limited information of cosmological relevance.
On the other hand, looking at the whole sky is often superfluous and waste computations, as most interactions are local.
A data and task dependent trade-off is to be found.
Note here that while global predictions use, by definition, the whole sky, dense predictions are not necessarily local.
